\documentclass{article}

% Toggle between anonymous submission and final version.
\newif\ifanonymized
\anonymizedtrue % Change to \anonymizedfalse for a named/final version.

% Paper metadata
\newcommand{\PaperTitle}{Verifier-Aligned Cover Complexity for CPWL Neural Network Verification}
\newcommand{\PaperShortTitle}{Verifier-Aligned Cover Complexity}
\newcommand{\PaperKeywords}{Computer-aided verification; neural network verification; piecewise-linear networks; proof complexity}

% Named author block for final/non-anonymous builds.
\newcommand{\CameraReadyAuthors}{%
  First Author \\
  Affiliation One \\
  Address \\
  \texttt{author1@example.com}
  \And
  Second Author \\
  Affiliation Two \\
  Address \\
  \texttt{author2@example.com}%
}

% Optional: enable compact numeric citations (common for space-constrained theory papers).
\PassOptionsToPackage{numbers,compress}{natbib}

\ifanonymized
  \usepackage{neurips_2026}
\else
  \usepackage[final]{neurips_2026}
\fi

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[hidelinks]{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{graphicx}

\graphicspath{{figures/}}

\title{\PaperTitle}

\author{%
\ifanonymized
  Anonymous Author(s) \\
  Affiliation \\
  Address \\
  \texttt{email}
\else
  \CameraReadyAuthors
\fi
}

\begin{document}

\maketitle

\begin{abstract}
Proof-producing verification is constrained by what a trusted checker can independently re-check, not by the internal search trace of a particular solver. For continuous piecewise-linear (CPWL) neural networks, however, complexity is still usually described through solver-centric or geometry-centric surrogates such as branch-and-bound nodes, mixed-integer encodings, or linear-region counts. We introduce \emph{Verifier-Aligned Cover Complexity} (VAC): the minimum number of verifier-visible local certified cells needed for an accepted proof of either exact semantics or a threshold property over a compact domain. For CPWL expression graphs built from unary affine post-composition, addition, scalar gates, and finite-max modules (equivalently, arbitrary affine nodes after normalization), we prove a guard-certified exactification theorem: a static compilation into guard-affine fragments yields finitely many comparison hyperplanes, and any split tree aligned with them turns every nonempty leaf into an accepted exact affine cell. Hence exact-side VAC is finite for this class. After fixing a verifier-facing language, we then show that cell-complete proof languages realize VAC up to constant factors, and that linear encodings preserve this scale up to constant factors as well, thereby separating semantic cover complexity from proof syntax and bit-level size. We also derive threshold lower bounds via incompatibility graphs, and prove that the explicit-dictionary decision problems \textsc{Exact-VAC-Dict} and \textsc{Thresh-VAC-Dict} are NP-complete while the corresponding optimization problems are hard to approximate within $(1-\varepsilon)\ln N$. VAC therefore isolates the language-relative semantic cover quantity underlying proof-producing CPWL verification, distinct from arrangement counts, solver traces, and syntax-specific encodings.
\end{abstract}

\section{Introduction}\label{sec:introduction}

Proof-producing verification lives at the boundary between search and trust. Search may involve SMT, MILP, branch-and-bound, abstract interpretation, or cutting-plane reasoning, but the trusted end of the pipeline only consumes a finite proof object that can be checked independently~\cite{Clarke1999ModelChecking,Necula1997PCC,Liu2021Survey,Bunel2018UnifiedView,Cordeiro2025PLChallenge}. For CPWL neural networks, however, complexity is still usually discussed through solver-centric or geometry-centric surrogates such as branch-and-bound nodes, SMT/MIP encodings, or linear-region counts~\cite{Katz2017Reluplex,Tjeng2019MIP,Bunel2020BranchBound,Montufar2014LinearRegions,Serra2018Counting}. These quantities matter algorithmically, but they are not the object that a proof-producing verifier ultimately accepts.

This mismatch has become more visible as proof-producing and checker-oriented neural network verification has matured~\cite{Isac2022ProofProduction,Fischer2022SharedCertificates,Desmartin2025ImandraChecker,Duong2025APTP}. Recent systems increasingly emphasize certificates, checker kernels, and reusable proof artifacts. Yet one question remains under-formalized: \emph{what is the verifier-facing semantic unit being counted?} If the trusted interface only sees local claims plus a global cover argument, then counting solver states or geometric regions is, at best, indirect.

Our starting point is therefore proof-theoretic. In our setting, the local unit consumed by the verifier is a \emph{local certified cell}: a verifier-visible cell together with a finite payload proving either exact affine semantics on that cell or a local threshold claim. Such cells need not coincide with arrangement regions. One accepted cell may merge several geometric pieces when one certificate suffices, while one geometric region may need further splitting because the chosen verifier-facing language cannot certify it directly.

We formalize this viewpoint by \emph{Verifier-Aligned Cover Complexity} (VAC), the minimum number of verifier-visible local certified cells needed for an accepted global proof. VAC is semantic rather than algorithmic, and language-relative by design: once the trusted interface is fixed, VAC counts the smallest accepted cover in that interface. It is therefore the natural quantity to compare against proof syntax and encoding size, not because languages are interchangeable, but precisely because they are \emph{not}.

For the CPWL expression graphs studied here---built from unary affine post-composition, addition, scalar gates, and finite-max modules over nonempty compact domains, with arbitrary affine fan-in normalized into that basis---we prove a guard-certified exactification theorem. A static guard compilation exposes finitely many comparison hyperplanes, and any split tree aligned with them turns every nonempty leaf into an accepted exact affine cell. This yields finite accepted exact covers for the whole class. We then separate three layers that are often conflated: the semantic cover quantity VAC, the leaf complexity of a chosen proof syntax (SCC), and its bit-level encoding size (DL). Finally, we give verifier-aligned lower bounds via incompatibility graphs and isolate explicit-dictionary cover selection as a genuinely hard combinatorial layer.

\paragraph{Contributions.}
The paper makes five contributions.
\begin{enumerate}
    \item \textbf{A structural theorem for exact-side finiteness.} For the CPWL class above, guard-certified exactification yields a finite accepted exact certified cover; equivalently, exact-side VAC is finite in the guard language.
    \item \textbf{A three-layer complexity decomposition.} After fixing a verifier-facing language, VAC captures semantic cover size, SCC captures syntactic realization cost, and DL captures bit-level encoding cost. Cell-complete proof languages realize VAC up to constant factors, and linear encodings preserve that scale up to constant factors.
    \item \textbf{Verifier-aligned upper and lower bounds.} Exactification gives accepted exact-side upper bounds, exact covers transfer to threshold covers when the threshold claim is true, and incompatibility graphs yield language-relative threshold lower bounds.
    \item \textbf{Selection-layer hardness.} The unrestricted explicit-dictionary decision problems \textsc{Exact-VAC-Dict} and \textsc{Thresh-VAC-Dict} are NP-complete, and the corresponding optimization problems inherit the classical logarithmic inapproximability barrier.
    \item \textbf{An artifact-backed empirical calibration.} The saved benchmark bundle calibrates exactification, threshold exposure, selection-stage compression, and realization-side payload scaling while carefully distinguishing accepted upper bounds, exposed dictionary optima, and engineering archive overhead.
\end{enumerate}

\paragraph{Scope and non-claims.}
The scope is intentionally narrow. We study CPWL expression graphs over nonempty compact domains. VAC is language-relative by design, so we do \emph{not} claim invariance across different verifier-facing interfaces. The hardness theorem is unconditional only for unrestricted explicit dictionaries; transfer to a concrete pipeline requires additional representability assumptions on the candidate families that the pipeline can expose. Empirically, Gold Tiny is the only exhaustive suite in the current artifact: outside it we report accepted upper bounds and dictionary-restricted optima over exposed candidate languages, not unrestricted ground-truth VAC values. On the threshold side in particular, the saved v3 artifact uses a bounded exposed threshold language, so threshold experiments must be read in that language-relative sense.

\paragraph{How to read the paper.}
The paper contains four logically distinct claim types. First, a \emph{structural} claim: guard-certified exactification yields finite accepted exact covers for the CPWL class considered here. Second, a \emph{realization} claim: once a verifier-facing language is fixed, proof languages and encodings track the same semantic cover scale up to fixed overhead. Third, a \emph{lower-bound} claim: incompatibility graphs witness threshold-side cover complexity inside the fixed language. Fourth, a \emph{combinatorial hardness} claim: explicit-dictionary cover selection is NP-complete and logarithmically hard to approximate. Section~\ref{sec:verifier-model} fixes the verifier interface, Section~\ref{sec:guard-exactification} proves finite exactification, Section~\ref{sec:vac-proof-languages} relates VAC to proof syntax and encoding, Section~\ref{sec:vac-bounds} gives upper and lower bounds, Section~\ref{sec:vac-minimization} isolates explicit-dictionary hardness, and Section~\ref{sec:results} summarizes the saved artifact. Appendix~\ref{app:trusted-substrate} gives the fuller trusted-task semantics, while Appendices~\ref{app:guard-rules}--\ref{app:dict-hardness} contain deferred proofs and auxiliary formalizations.

\section{Verifier-Aligned Certified Covers and Main Results}\label{sec:verifier-model}

This section fixes the minimal API between an untrusted search procedure and the trusted checker. Everything counted later is defined at this interface. Appendix~\ref{app:trusted-substrate} develops the fuller task semantics; here we keep only the load-bearing objects.

Fix a nonempty compact domain $D\subseteq\mathbb R^n$ and a target function $g:D\to\mathbb R$. Individual verifier-visible cells $C(\gamma)$ may still be empty; such cells are discharged by \textsc{Feas}, and whenever an affine infimum over an empty cell appears we use the convention $\inf_{x\in\varnothing} a(x)=+\infty$.

A \emph{verifier-aligned cell language} $\mathfrak L$ consists of finite descriptions $\gamma$ with verifier-facing set description $\mathfrak D(\gamma)$ and semantic cell
\[
C(\gamma):=[\![\mathfrak D(\gamma)]\!]\subseteq D.
\]

\paragraph{Trusted kernel.}
The kernel exposes only four primitives:
\begin{itemize}
    \item \textsc{Feas}, which either finds $x\in[\![\mathfrak D]\!]$ or proves emptiness;
    \item \textsc{AffOpt}, which certifies $\inf_{x\in[\![\mathfrak D]\!]}(w^\top x+b)$ for an affine tag;
    \item \textsc{Eval}, which returns a certified value $g(x)$ at a queried point; and
    \item \textsc{Check}, which deterministically validates finite symbolic payloads such as containments, local equalities or threshold claims, and finite cover proofs.
\end{itemize}
Search traces, branching heuristics, and solver internals are never trusted.

\paragraph{Certified cells and covers.}
For threshold $t\in\mathbb R$, an \emph{exact certified cell} is $(\gamma,a,\pi^{=})$, where $a$ is affine and the accepted payload proves $g(x)=a(x)$ for all $x\in C(\gamma)$; a \emph{threshold certified cell} is $(\gamma,\pi^{\ge})$, where the accepted payload proves $g(x)\ge t$ on $C(\gamma)$. Finite families of such cells become exact or threshold certified covers once \textsc{Check} accepts a global cover payload for
\[
D\subseteq \bigcup_i C(\gamma_i).
\]
The carrier is the same on both sides; only the local statement changes.

\paragraph{VAC.}
We now count the minimum number of verifier-visible local proof units needed for an accepted global proof:
\[
\mathrm{VAC}^{=}_{\mathfrak L}(g;D) := \min\{\,|\mathcal C^{=}|:\ \mathcal C^{=}\text{ is an exact certified cover of }(g,D)\,\},
\]
\[
\mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t) := \min\{\,|\mathcal C^{\ge}|:\ \mathcal C^{\ge}\text{ is a threshold certified cover of }(g,D,t)\,\},
\]
with value $+\infty$ if no finite accepted cover exists. VAC therefore counts verifier-visible local certified cells, not runtime, search depth, branch-and-bound nodes, or arrangement regions.

\paragraph{Proposition (basic soundness contract).}
Accepted exact cells state correct affine equalities on their cells, and accepted threshold cells state correct local threshold claims. Consequently, if $\mathcal C^{=} = \{(\gamma_i,a_i,\pi_i^{=})\}_{i=1}^N$ is an exact certified cover, then
\[
\inf_{x\in D} g(x)= \min_{1\le i\le N}\ \inf_{x\in C(\gamma_i)} a_i(x),
\]
with the convention $\inf_{x\in\varnothing} a_i(x)=+\infty$; and if $\mathcal C^{\ge}$ is a threshold certified cover, then $g(x)\ge t$ for all $x\in D$.

\paragraph{Main theorem package.}
The rest of the paper proves four facts.
\begin{enumerate}
    \item \textbf{Finite exactification for CPWL graphs.} Section~\ref{sec:guard-exactification} constructs a guard language $\mathfrak L_{\mathrm{grd}}$ for the allowed CPWL expression graphs and proves $\mathrm{VAC}^{=}_{\mathfrak L_{\mathrm{grd}}}(g;D)<\infty$ by compiling the network into finitely many guard-affine fragments and aligning a split tree with the resulting comparison hyperplanes.

    \item \textbf{VAC, proof languages, and encoding overhead.} Section~\ref{sec:vac-proof-languages} shows that any $c$-cell-complete proof language $\mathcal P$ over $\mathfrak L$ satisfies
    \[
    \mathrm{VAC}^{\star}_{\mathfrak L} \le \mathrm{SCC}^{\star}_{\mathcal P} \le c\,\mathrm{VAC}^{\star}_{\mathfrak L},
    \qquad \star\in\{=,\ge\},
    \]
    and under linear encodings $\mathrm{DL}^{\star}_{\mathcal P}$ is linearly equivalent to $\mathrm{SCC}^{\star}_{\mathcal P}$; see Appendix~\ref{app:proof-language-dl}.

    \item \textbf{Verifier-aligned upper and lower bounds.} Section~\ref{sec:vac-bounds} gives exact-side upper bounds from accepted exact covers and threshold-side lower bounds
    \[
    \mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t)\ge \chi(G_t^{\mathrm{inc}}(X))
    \]
    from incompatibility graphs. When $\inf_D g\ge t$, exact certified covers also transfer to threshold certified covers of no larger size.

    \item \textbf{Selection-layer hardness.} Section~\ref{sec:vac-minimization} proves NP-completeness for the decision problems \textsc{Exact-VAC-Dict} and \textsc{Thresh-VAC-Dict}, together with $(1-\varepsilon)\ln N$ hardness for the corresponding dictionary-restricted optimization problems; Appendix~\ref{app:dict-hardness} gives the full reductions.
\end{enumerate}

Conceptually, this section is only the paper's API. Later sections construct objects in this API, count them, upper-bound them, lower-bound them, and show that explicit selection among them is hard.

\section{Guard-Certified Exactification and Finite Exact Covers for CPWL Networks}\label{sec:guard-exactification}

This section proves the structural theorem behind exact-side finiteness. The proof has three steps: compile local guard-affine fragments, align a split tree with the finite set of resulting comparison hyperplanes, and attach exact leaf certificates to the live leaves.

We consider finite DAG expression graphs whose scalar nodes are built from four modules: unary affine post-composition nodes, addition nodes, scalar gates such as ReLU/LeakyReLU/PReLU/Abs, and finite-max nodes. Any finite-input affine map is first normalized into unary affine post-compositions plus additions, so this choice of primitives is without loss of generality for the structural arguments below. Vector outputs are handled coordinatewise. Let $u_g$ be the output node and write $g(x)=[\![u_g]\!](x)$ on nonempty compact $D$.

\paragraph{Guard language.}
The exact-side language $\mathfrak L_{\mathrm{grd}}$ uses finite sets of affine comparison literals $h(x)\le 0$ and $h(x)\ge 0$. A guard set $S$ denotes
\[
D_S:=D\cap \bigcap_{\lambda\in S}\lambda.
\]
For a scalar node $u$, a \emph{certified guard-affine fragment} is a triple $(S,a,\Pi)$, where $a$ is affine and $\Pi$ is accepted as a proof that $[\![u]\!](x)=a(x)$ for all $x\in D_S$. A finite family $\mathcal F_u=\{(S_r,a_r,\Pi_r)\}$ is \emph{certified faithful} if the cells $D_{S_r}$ cover $D$ and each payload proves the corresponding local affine equality.

\paragraph{Static compilation.}
Compilation proceeds by topological induction using only four finite rules: affine propagation, common-refinement addition, gate branching via an affine winner comparison, and finite-max with explicit winner comparisons. Each rule combines already accepted child payloads with finitely many affine guards, so faithfulness is preserved.

\paragraph{Theorem (static guard-certified compilation).}
Every scalar node $u$ in an allowed CPWL expression graph admits a finite certified faithful family $\mathcal F_u$. Applied to the output node, this yields a finite family $\mathcal F_g$ and therefore a finite comparison-hyperplane set $\mathcal H^\star$ produced by compilation, whose leafwise comparison information suffices to freeze the relevant gate branches and max comparisons. Input coordinates and constants give trivial one-fragment families, and the four rules preserve certified faithfulness inductively. Appendix~\ref{app:guard-rules} gives the full rule-by-rule proof.

\paragraph{Aligned leaves.}
A split tree recursively branches on literals from $\mathcal H^\star$; each leaf $v$ carries a path guard set $S(v)$ and domain $D_v:=D_{S(v)}$. The tree is \emph{aligned} with $\mathcal H^\star$ if, for every leaf and every $h\in\mathcal H^\star$, \textsc{Check} accepts at least one side claim $D_v\subseteq\{h\le 0\}$ or $D_v\subseteq\{h\ge 0\}$. Alignment freezes every gate branch choice and max-winner comparison leafwise.

\paragraph{Lemma (aligned nonempty leaves are exact).}
If $T$ is aligned with $\mathcal H^\star$ and $D_v\neq\varnothing$, then some fragment $(S,a,\Pi)\in\mathcal F_g$ is valid on all of $D_v$; restricting $\Pi$ along the accepted containment $D_v\subseteq D_S$ yields an exact certified cell $(S(v),a_v,\pi_v^{=})$ with $g(x)=a_v(x)$ for all $x\in D_v$. Thus every nonempty aligned leaf is an atomic exact cell in $\mathfrak L_{\mathrm{grd}}$. The proof is an aligned-leaf induction showing that the leaf signature fixes every dynamic choice in the compilation; see Appendix~\ref{app:exactification}.

\paragraph{Global cover payload.}
A finite split tree also carries a canonical cover payload: every internal node is covered by its two closed-halfspace children, and infeasible leaves can be removed once \textsc{Feas} proves emptiness. Hence exactification needs only a finite fully aligned tree and the accepted local payloads attached to its live leaves.

\paragraph{Theorem (guard-certified exactification).}
For the output of any allowed CPWL expression graph over compact $D$, there exists a finite split tree $T^\star$ aligned with $\mathcal H^\star$ such that every nonempty leaf yields an exact certified cell and the live leaves, together with the split-tree cover payload, form an accepted finite exact certified cover of $g$ in $\mathfrak L_{\mathrm{grd}}$. Consequently,
\[
\mathrm{VAC}^{=}_{\mathfrak L_{\mathrm{grd}}}(g;D)<+\infty.
\]
The theorem is structural, not optimal: $|\operatorname{Live}(T^\star)|$ is one verifier-accepted upper bound, not necessarily the minimum.

\section{VAC, Proof Languages, and Realization Overhead}\label{sec:vac-proof-languages}

Section~\ref{sec:guard-exactification} gives one accepted exact cover. We now separate three layers that are often conflated: semantic cover size, concrete proof syntax, and bit-level encoding size.

Fix a verifier-facing language $\mathfrak L$. A \emph{proof language} $\mathcal P$ over $\mathfrak L$ consists of finite proof objects whose leaves denote certified cells from $\mathfrak L$ and whose global payload certifies that the leaves jointly form the required cover. This includes flat leaf tables, split trees, and more structured proof objects with the same verifier-visible leaf semantics.

\paragraph{Syntactic and bit-level measures.}
The minimum leaf count of an accepted proof object defines syntactic certificate complexity:
\[
\mathrm{SCC}^{=}_{\mathcal P}(g;D)
:=
\min\{\,|\operatorname{Leaves}(P)|:\ P\in\mathcal P\text{ is an accepted exact proof of }(g,D)\,\},
\]
\[
\mathrm{SCC}^{\ge}_{\mathcal P}(g;D,t)
:=
\min\{\,|\operatorname{Leaves}(P)|:\ P\in\mathcal P\text{ is an accepted threshold proof of }(g,D,t)\,\}.
\]
Fixing a binary encoding analogously gives $\mathrm{DL}^{=}_{\mathcal P}$ and $\mathrm{DL}^{\ge}_{\mathcal P}$. The canonical flat leaf-table language $\mathcal P_{\mathrm{tab}}(\mathfrak L)$ realizes VAC exactly, because a semantic certified cover is already such a table plus one cover payload.

\paragraph{Cell-complete languages.}
A proof language is $c$-cell-complete if every accepted proof object extracts to a certified cover with no more cells than leaves, and every certified cover of size $N$ compiles into an accepted proof object with at most $cN$ leaves. The exact and threshold sides are defined separately but identically.

\paragraph{Theorem (VAC/SCC equivalence under cell completeness).}
If $\mathcal P$ is exact-side or threshold-side $c$-cell-complete over $\mathfrak L$, then for $\star\in\{=,\ge\}$,
\[
\mathrm{VAC}^{\star}_{\mathfrak L}
\le
\mathrm{SCC}^{\star}_{\mathcal P}
\le
c\,\mathrm{VAC}^{\star}_{\mathfrak L}.
\]
The left inequality extracts a semantic cover from any accepted proof object; the right inequality compiles an optimal semantic cover back into the chosen syntax with at most $c$ times as many leaves. Under the linear-encoding assumptions formalized in Appendix~\ref{app:proof-language-dl}, $\mathrm{DL}^{\star}_{\mathcal P}$ is likewise linearly equivalent to $\mathrm{SCC}^{\star}_{\mathcal P}$.

The point is not cross-language invariance. The point is layer separation: after fixing $\mathfrak L$, VAC measures the semantic cover quantity, SCC measures syntactic realization cost, and DL measures encoding cost.

\paragraph{Arrangement count as a restricted degeneration.}\label{subsec:arrangement-degeneration}
Arrangement geometry reappears only under extra exact-side assumptions. If the maximal connected exact-affine regions $U_1,\dots,U_M$ of $g$ are each realizable as one connected exact certified cell of $\mathfrak L$, and each region contains a point not lying in any other region, then the minimum number of connected exact certified cells is exactly $M$. The upper bound comes from one cell per region; the lower bound uses the region-exclusive witness points, which no connected exact cell can cover across two different maximal regions. Outside this narrow regime arrangement count is not, in general, the semantic cover quantity: a verifier-admissible cell may refine or merge geometric pieces, and threshold proofs need not recover exact affine regions at all. Appendix~\ref{app:proof-language-dl} expands the proof-language formalization and encoding details.

\section{Upper and Lower Bounds for Verifier-Aligned Covers}\label{sec:vac-bounds}

This section packages the quantitative consequences of the previous two sections. Exactification supplies accepted exact covers for the CPWL class studied here, and the proof-language formalization shows how those semantic quantities transfer to syntactic realizations up to fixed constants. We now state the first bounds directly at the VAC layer.

\subsection{Exact-Side Upper Bounds from Finite Exactification}

Fix an allowed CPWL expression graph over nonempty compact domain $D$, let $g$ be its output, and let
\[
\mathcal H^\star
\]
be the finite comparison-hyperplane family produced in Section~\ref{sec:guard-exactification}. If $T$ is any split tree aligned with $\mathcal H^\star$, write
\[
\operatorname{Live}(T)
:=
\{v\in \operatorname{Leaves}(T): D_v\neq\varnothing\}.
\]
By the exactification theorem, each $v\in \operatorname{Live}(T)$ carries an exact certified cell
\[
\bigl(S(v),a_v,\pi_v^{=}\bigr)
\]
in the guard language $\mathfrak L_{\mathrm{grd}}$, and Section~\ref{sec:guard-exactification} also provides the finite global cover payload certifying that the live leaves cover $D$.

\paragraph{Theorem (Exact-side upper bound from finite exactification).}
Let $T$ be any split tree aligned with $\mathcal H^\star$. Then the family
\[
\Bigl\{\bigl(S(v),a_v,\pi_v^{=}\bigr): v\in \operatorname{Live}(T)\Bigr\}
\]
together with the split-tree cover payload is an exact certified cover of $D$. Consequently,
\[
\mathrm{VAC}^{=}_{\mathfrak L_{\mathrm{grd}}}(g;D)
\le
|\operatorname{Live}(T)|.
\]
In particular,
\[
\mathrm{VAC}^{=}_{\mathfrak L_{\mathrm{grd}}}(g;D)<+\infty.
\]

\smallskip
\noindent\emph{Proof sketch.}
Section~\ref{sec:guard-exactification} proves three facts simultaneously: every aligned nonempty leaf is an accepted exact certified cell; every aligned empty leaf is discharged by infeasibility; and the split tree itself carries a finite global cover payload. Therefore the live leaves are not merely a semantic partition into exact pieces; they already form an accepted exact certified cover in the sense of Section~\ref{sec:verifier-model}. VAC is the minimum size of such a cover, so it cannot exceed the number of live leaves.

The theorem is a finiteness theorem, not an optimality theorem. The right-hand side counts one accepted exactification cover, not arrangement regions of the underlying function geometry. A smaller verifier-accepted cover may still exist because exact certified cells need not coincide one-to-one with connected exact-affine regions.

\subsection{Exact-to-Threshold Transfer}

The exact-side theorem gives finite accepted exact covers unconditionally for the CPWL class considered here. The threshold side is different: a finite threshold certified cover can exist only when the threshold claim itself is true. The transfer from exact proofs to threshold proofs is therefore conditional on semantic truth.

\paragraph{Proposition (Exact-to-threshold transfer).}
Fix any verifier-facing cell language $\mathfrak L$ and threshold $t\in\mathbb R$. If
\[
\inf_{x\in D} g(x)\ge t,
\]
then every exact certified cover of $(g,D)$ with $N$ cells induces a threshold certified cover of $(g,D,t)$ with the same $N$ cell descriptions. Consequently,
\[
\mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t)
\le
\mathrm{VAC}^{=}_{\mathfrak L}(g;D).
\]

\smallskip
\noindent\emph{Proof sketch.}
Take one exact certified cell
\[
(\gamma,a,\pi^{=}).
\]
Because
\[
g(x)=a(x)\qquad \forall x\in C(\gamma),
\]
and because the global assumption gives $\inf_D g\ge t$, we have
\[
\inf_{x\in C(\gamma)} a(x)
=
\inf_{x\in C(\gamma)} g(x)
\ge t.
\]
Calling \textsc{AffOpt} on the verifier-facing description $\mathfrak D(\gamma)$ and the affine tag $a$ yields a finite certificate of this local affine infimum; no optimizer witness is needed. Concatenating that local optimization certificate with the already accepted exact payload gives a finite threshold payload on the same cell description $\gamma$. Repeating this cell by cell converts any exact certified cover into a threshold certified cover without changing its cardinality.

\paragraph{Corollary (Threshold-side finiteness is conditional).}
For every verifier-facing language $\mathfrak L$,
\[
\inf_{x\in D} g(x)<t
\quad\Longrightarrow\quad
\mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t)=+\infty.
\]
For the CPWL guard language,
\[
\inf_{x\in D} g(x)\ge t
\quad\Longrightarrow\quad
\mathrm{VAC}^{\ge}_{\mathfrak L_{\mathrm{grd}}}(g;D,t)
\le
\mathrm{VAC}^{=}_{\mathfrak L_{\mathrm{grd}}}(g;D)
\le
|\operatorname{Live}(T)|
\]
for every split tree $T$ aligned with $\mathcal H^\star$.

\smallskip
\noindent\emph{Reason.}
If a finite threshold certified cover existed while $\inf_D g<t$, then the soundness contract from Section~\ref{sec:verifier-model} would force $g(x)\ge t$ for all $x\in D$, a contradiction. When the threshold claim is true, the proposition applies to the exactification cover from the previous subsection.

Thus exactification always yields exact-side finiteness, whereas threshold-side finiteness is a semantic consequence of truth. This is why threshold VAC is not merely ``exact VAC with inequalities'': it measures a different proof task.

\subsection{Threshold Lower Bounds via Incompatibility Graphs}

Upper bounds show that verifier-aligned covers exist. To prove that they sometimes must be large, we move to a clean threshold-side lower-bound template. The question is purely semantic: which witness points can one accepted local threshold cell certify simultaneously?

Fix a verifier-facing language $\mathfrak L$ and threshold $t$.

\paragraph{Definition (Threshold incompatibility graph).}
Let
\[
X=\{x_1,\dots,x_m\}\subseteq D
\]
be a finite witness set. Two points $x_i,x_j\in X$ are \emph{threshold-compatible} if there exists a threshold certified cell
\[
(\gamma,\pi^{\ge})
\]
such that
\[
x_i,x_j\in C(\gamma).
\]
If no such threshold certified cell exists, they are \emph{threshold-incompatible}. The corresponding incompatibility graph is the graph
\[
G_t^{\mathrm{inc}}(X)
\]
on vertex set $X$ with edge relation
\[
\{x_i,x_j\}\in E\bigl(G_t^{\mathrm{inc}}(X)\bigr)
\iff
x_i\text{ and }x_j\text{ are threshold-incompatible}.
\]
This graph is language-relative and verifier-aligned: it depends only on which pairs of points can be handled together by one accepted local threshold certificate.

\paragraph{Observation.}
Every threshold certified cell induces an independent set in $G_t^{\mathrm{inc}}(X)$. Indeed, if two vertices contained in the same accepted threshold cell were adjacent, then by definition they would be threshold-incompatible, which is impossible.

\paragraph{Theorem (Chromatic-number lower bound).}
For every finite witness set $X\subseteq D$,
\[
\mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t)
\ge
\chi\bigl(G_t^{\mathrm{inc}}(X)\bigr),
\]
where $\chi$ denotes chromatic number.

\smallskip
\noindent\emph{Proof sketch.}
Take any threshold certified cover
\[
\mathcal C^{\ge}=\{(\gamma_i,\pi_i^{\ge})\}_{i=1}^N.
\]
Each cell contributes an independent vertex set
\[
I_i:=X\cap C(\gamma_i)
\]
of the incompatibility graph. Because the cells cover $D$, the sets $I_1,\dots,I_N$ cover every vertex of $X$. Hence the graph is covered by $N$ independent sets. The minimum number of independent sets needed to cover the vertices of a finite graph is exactly its chromatic number, so every threshold certified cover must have size at least $\chi(G_t^{\mathrm{inc}}(X))$. Minimizing over all covers gives the claim.

Two corollaries are especially convenient.

\paragraph{Corollary (Clique lower bound).}
For every finite witness set $X\subseteq D$,
\[
\mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t)
\ge
\omega\bigl(G_t^{\mathrm{inc}}(X)\bigr),
\]
where $\omega$ is the clique number.

\paragraph{Corollary (Pairwise incompatibility).}
If there exist points $x_1,\dots,x_m\in D$ that are pairwise threshold-incompatible, then
\[
\mathrm{VAC}^{\ge}_{\mathfrak L}(g;D,t)\ge m.
\]

These lower bounds are as verifier-aligned as the upper bounds from exactification. They do not count search nodes, branch-and-bound iterations, or linear regions. They record only which witness points one accepted local threshold certificate can or cannot handle together. In that sense they are proof-complexity lower bounds at the fixed semantic cover layer.

Taken together, Sections~\ref{sec:guard-exactification}, \ref{sec:vac-proof-languages}, and the present section give the core quantitative picture of the paper: exactification supplies accepted finite exact covers, threshold incompatibility supplies language-relative lower bounds, and VAC is the fixed-language semantic cover quantity that both sides meet on. Section~\ref{sec:vac-minimization} then asks the complementary algorithmic question: once an explicit finite pool of candidate verifier-visible cells has been exposed, how hard is it to find the smallest accepted subcover?

\section{Selection-Layer Complexity for Explicit Dictionaries}\label{sec:vac-minimization}

The previous sections are semantic. We now isolate the residual combinatorial problem that remains once a finite family of verifier-visible candidate cells has already been exposed: choose the smallest accepted subcover. Full reductions are in Appendix~\ref{app:dict-hardness}.

\paragraph{Dictionary-restricted problems.}
Work over an explicit finite rational point domain $U=\{u_1,\dots,u_N\}\subseteq\mathbb Q^n$ with explicit values $g(u_i)\in\mathbb Q$. An exact dictionary entry is $(A_j,a_j)$ with $A_j\subseteq U$ and affine $a_j$; it is valid when $g(u)=a_j(u)$ for all $u\in A_j$. \textsc{Exact-VAC-Dict} asks whether at most $k$ valid entries from a dictionary $\mathcal D_=$ cover $U$, and $\operatorname{OPT}_{=}(U,\mathcal D_=)$ denotes the minimum size of such a cover. For rational threshold $t$, \textsc{Thresh-VAC-Dict} is defined analogously using valid threshold entries $A_j\subseteq U$ satisfying $g(u)\ge t$ for all $u\in A_j$, with optimum $\operatorname{OPT}_{\ge}(U,\mathcal D_{\ge},t)$.

These problems isolate only the \emph{selection layer}: no continuous search or candidate generation remains; all residual difficulty is combinatorial.

\paragraph{Theorem (NP-completeness of explicit-dictionary VAC selection).}
The decision problems \textsc{Exact-VAC-Dict} and \textsc{Thresh-VAC-Dict} are NP-complete. Membership in NP is immediate because a proposed subcover can be checked by finite inspection. For NP-hardness, reduce from \textsc{Set Cover}: encode each universe element as a rational point, set $g\equiv 0$, and use one dictionary entry per set, with affine tag $a_j\equiv 0$ on the exact side and threshold $t=0$ on the threshold side. Valid subcovers of the dictionary are then exactly set covers, and the optimum value is preserved.

\paragraph{Corollary (Logarithmic inapproximability of explicit-dictionary VAC minimization).}
For every fixed $\varepsilon\in(0,1)$, unless $\mathrm P=\mathrm{NP}$, neither corresponding optimization problem $\operatorname{OPT}_{=}(U,\mathcal D_=)$ nor $\operatorname{OPT}_{\ge}(U,\mathcal D_{\ge},t)$ admits a polynomial-time approximation within factor $(1-\varepsilon)\ln N$. The classical logarithmic barrier for \textsc{Set Cover} transfers directly through the optimum-preserving reduction.

The conclusion is deliberately scoped. Appendix~\ref{app:dict-hardness} proves hardness for the unrestricted explicit-dictionary selection problem. A concrete proof pipeline inherits this hardness only if the class of candidate families it can expose is rich enough to realize hard set-cover instances, formalized in Theorem~G.15. Absent such a representability assumption, no generic hardness transfer follows from the present reduction alone. Candidate generation and cover selection are therefore distinct sources of difficulty.

\section{Experimental Results}\label{sec:results}

This section summarizes the saved v3 artifact; Appendix~\ref{app:experiments} provides the full figures and per-family tables. The reporting principle follows the theory developed above: we compare verifier-facing cover quantities and their realizations, not solver-trace proxies.

Gold Tiny is the only exhaustive suite in the current artifact. Outside Gold Tiny we therefore do \emph{not} claim unrestricted ground-truth VAC values. On the exact side we report accepted exactification upper bounds $U_{\mathrm{eq}}^{\mathrm{align}}$ together with dictionary-restricted optima over the exposed exact candidate language. On the threshold side we report transfer upper bounds $U_{\ge}^{\mathrm{transfer}}$, dictionary-restricted optima over the bounded exposed threshold language implemented by the v3 pipeline, and certified exposed lower bounds. Realization metrics are split into compact/core payloads and full archive bundles.

\paragraph{Exact-side calibration and scaling.}
The exact-side results support the structural theorem. All 126 exact-side instances complete, and $U_{\mathrm{eq}}^{\mathrm{align}}$ grows with the number $H=|\mathcal H^\star|$ of compiled comparison hyperplanes while staying far below the crude $2^H$ envelope (Appendix~\ref{app:experiments}, Fig.~\ref{fig:appI-exact-scaling}). The median accepted exact upper bounds are $32.5$, $34.0$, and $33.0$ on Family A, Family B, and ThresholdStress, compared with $2.0$ on the exhaustive Gold Tiny suite. Gold Tiny also plays a calibration role: in five of the six tiny cases, arrangement count, true guard-language exact VAC, exposed exact optimum, and accepted exactification upper bound coincide, while the designed maxout case separates them. This is exactly the semantic distinction the paper predicts: arrangement geometry, semantic cover size, and constructive upper bounds may agree in simple cases, but they need not agree in general. Exact compression is usually tight: the median ratio $U_{\mathrm{eq}}^{\mathrm{align}} / \mathrm{OPT}_{\mathrm{eq}}^{\mathrm{exposed}}$ is $1.0\times$ on Family A, Family B, Gold Tiny, and ThresholdStress. The main exception is SelectionStress, whose median compression factor is $12.5\times$, confirming that candidate exposure and semantic optimum can separate substantially even when the constructive exactification bound remains moderate.

\paragraph{Threshold-side behavior and scope.}
The threshold results match the theory but do not support a universal ``threshold is easier'' slogan. On all 126 false controls the artifact finds no finite accepted threshold cover. On all 126 canonical true-threshold tasks, the transfer upper bound dominates the exposed threshold optimum, exactly as required by the exact-to-threshold transfer theorem. On the curated threshold subset, enlarging the exposed threshold language from $D_{\mathrm{transfer}}$ to $+D_{\mathrm{drop}}$ reduces the median exposed optimum from $38.0$ to $19.5$, while the median exposed lower bound stays at $16.0$. In most benchmark families the exposed threshold optimum is smaller than the exposed exact optimum, but this is not universal: the only 12 instances with $\mathrm{OPT}_{\ge}^{\mathrm{exposed}} > \mathrm{OPT}_{=}^{\mathrm{exposed}}$ all come from SelectionStress. We therefore state threshold conclusions in the same language-relative form used throughout the repository and appendix, rather than as a blanket claim of threshold dominance.

\paragraph{Selection exposure.}
The selection experiment isolates what happens after a finite candidate family has already been exposed. On the curated explicit-candidate subset, the selection dictionary grows from a median of $40.5$ candidates at $D_{\mathrm{leaf}}$ to $586.0$ at $+D_{\mathrm{drop}}$, while the median exposed exact optimum collapses from $40.5$ to $3.0$. This is direct evidence of selection-layer compression after candidate exposure. In other words, the exposed candidate family can become much richer before any theorem-level hardness statement is invoked.

\paragraph{Realization-side payloads.}
Compact/core proof size is almost perfectly linear in proof-cell count (Pearson correlation $0.995$, fitted slope $\approx 206$ bytes per cell), whereas archive bundles are much larger and strongly family-dependent, with median archive-to-core ratios ranging from $5.1\times$ on Gold Tiny to $251.7\times$ on SelectionStress. The artifact should therefore be read in two layers: compact/core payloads are the natural SCC/DL-facing realization object, whereas archives reflect engineering overhead. The saved bundle's compact/core checker replay stays below one millisecond, and exactification generation stays below $5.8$ seconds on every shipped instance. We report the checker numbers as realization-side artifact measurements for the saved compact/core payloads, not as a benchmark of a separately reconstructed external checker.

Overall, the experiments support three empirical claims: exactification yields observable finite covers at moderate constructive cost, threshold-side and selection-side exposure can materially change the optimum before any hardness statement is applied, and compact/core realizations---not full archives---are the correct object to compare at the SCC/DL-facing layer.

\section{Discussion and Conclusion}

This paper resolves one specific proof-complexity question for proof-producing CPWL verification. For the CPWL expression graphs studied here, guard-certified exactification yields finite accepted exact covers and therefore finite exact-side VAC. After fixing a verifier-facing language, the paper separates the semantic cover quantity from its syntactic and bit-level realizations, gives verifier-aligned upper and lower bounds, and shows that explicit-dictionary cover minimization already contains a genuine combinatorial hardness source.

The main take-away is layered rather than absolute. Arrangement count, solver traces, proof syntax, encoded proof size, and engineering archive size are all meaningful quantities, but they are \emph{not} the same quantity. Once the verifier-facing language is fixed, VAC identifies the semantic cover quantity being realized; SCC and DL then record realization and encoding overhead relative to that semantic layer.

The paper also makes its own limits explicit. The structural theorem is proved only for the CPWL graph class studied here and over nonempty compact domains. VAC is language-relative by design. The unconditional hardness theorem applies to unrestricted explicit dictionaries, with transfer to concrete pipelines requiring representability assumptions. Experimentally, Gold Tiny is the only exhaustive suite, and threshold-side exposed quantities are measured in the bounded threshold language implemented by the artifact. On the realization side, the compact/core checker metric should be read as saved-payload replay cost rather than as a standalone artifact-only checking theorem.

These limitations point to natural next steps: broader verifier-facing languages, tighter threshold-side constructive procedures, and standalone serialized proof formats whose checking cost can be measured independently of the in-memory exactification context. Even within the present scope, however, the conceptual message is stable: proof-producing CPWL verification has a semantic cover layer that deserves to be named and analyzed on its own. VAC is that layer.

% Uncomment for the final version if needed.
% \begin{ack}
% Replace this paragraph with acknowledgments, funding disclosure, and competing-interest disclosure.
% \end{ack}

\newpage
\bibliographystyle{plainnat}
\bibliography{refs}

\newpage
\appendix
\input{appendix/Appendix_A}
\input{appendix/Appendix_B}
\input{appendix/Appendix_C}
\input{appendix/Appendix_D}
\input{appendix/Appendix_E}
\input{appendix/Appendix_F}
\input{appendix/Appendix_G}
\input{appendix/Appendix_H}
\input{appendix/Appendix_I}

\newpage

\end{document}
